\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-09 papers/jss/article.Rnw
Oops check in the Rnw rather than the generated tex file.  Also, add section \labels and add a roadmap section at the end of the introduction that highlights the main sections.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@730 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@

\abstract{Modern data collection and analysis pipelines often involve
 a sophisticated mix of applications written in general purpose and
 specialized programming languages.  Protocol Buffers are a popular
 method of serializing structured data between applications---while remaining
 independent of programming languages or operating system.  The
 \CRANpkg{RProtoBuf} package provides a complete interface between this
 library and the R environment for statistical computing.
 %TODO(ms) keep it less than 150 words.
}

@@

% TODO(mstokely): If we go to JSS, include a larger paragraph here
% referencing each numbered section.  I don't like these generally,
% but its useful for this paper I think because we have a boring bit
% in the middle (full class/method details) and interesting
% applications at the end.
This article describes the basics of Google's Protocol Buffers through
an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
several common use cases for protocol buffers in data analysis.

\section{Protocol Buffers}

FIXME Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

@@

\section{Basic Usage: Messages and Descriptors}

@@

\section{Under the hood: S4 Classes, Methods, and Pseudo Methods}

@@

\section{Type Coercion}

@@

\section{Evaluation: data.frame to Protocol Buffer Serialization}

@@

\section{Application: Data Interchange in Web Services}

@@

\section{Application: Distributed Data Collection with MapReduce}

TODO(mstokely): Make this better.

Many large data sets in fields such as particle physics and
information processing are stored in binned or histogram form in order
to reduce the data storage requirements
\citep{scott2009multivariate}. Protocol Buffers make a particularly
good data transport format in distributed MapReduces environments
where large numbers of computers process a large data set for analysis.

There are two common patterns for generating histograms of large data
sets with MapReduce.  In the first method, each mapper task can
generate a histogram over a subset of the data that is has been
assigned, and then the histograms of each mapper are sent to one or
more reducer tasks to merge.

In the second method, each mapper rounds a data point to a bucket
width and outputs that bucket as a key and '1' as a value.  Reducers
then sum up all of the values with the same key and output to a data store.

In both methods, the mapper tasks must choose identical
bucket boundaries even though they are analyzing disjoint parts of the
input set that may cover different ranges, or we must implement
multiple phases.

>>>>

\documentclass[article]{jss}
\usepackage{booktabs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}

%% almost as usual
\author{Dirk Eddelbuettel\\Debian and R Projects \And 
        Murray Stokely\\Google, Inc}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Modern data collection and analysis pipelines often involve
a sophisticated mix of applications written in general purpose and
specialized programming languages.  Protocol Buffers are a popular
method of serializing structured data between applications---while remaining
independent of programming languages or operating system.  The
\CRANpkg{RProtoBuf} package provides a complete interface between this
library and the R environment for statistical computing.
%TODO(ms) keep it less than 150 words.
}
\Keywords{r, protocol buffers, serialization, cross-platform}
\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Dirk Eddelbuettel\\
  \\
  Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94040\\
  USA\\
  E-mail: \email{mstokely@google.com}\\
  URL: \url{http://www.stokely.org/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

@@

@@

% TODO(mstokely): If we go to JSS, include a larger paragraph here
% referencing each numbered section.  I don't like these generally,
% but its useful for this paper I think because we have a boring bit
% in the middle (full class/method details) and interesting
% applications at the end.

Section~\ref{sec:protobuf} provides a general overview of Protocol
Buffers.  Section~\ref{sec:rprotobuf-basic} describes the interactive
R interface provided by \CRANpkg{RProtoBuf} and introduces the two
main abstractions: \emph{Messages} and \emph{Descriptors}.
Section~\ref{sec:rprotobuf-classes} describes the implementation
details of the main S4 classes making up this package.
Section~\ref{sec:types} describes the challenges of type coercion
between R and other languages.  Section~\ref{sec:evaluation}
introduces a general R language schema for serializing arbitrary R
objects and evaluates it against R's built-in serialization.
Sections~\label{sec:opencpu} and \label{sec:mapreduce} provide
real-world use cases of \CRANpkg{RProtoBuf} in web service and
MapReduce environments, respectively.

%This article describes the basics of Google's Protocol Buffers through
%an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
%basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
%several common use cases for protocol buffers in data analysis.

\section{Protocol Buffers}
\label{sec:protobuf}

%FIXME Introductory section which may include references in parentheses
%\citep{R}, or cite a reference such as \citet{R} in the text.

@@

\section{Basic Usage: Messages and Descriptors}
\label{sec:rprotobuf-basic}

@@

\section{Under the hood: S4 Classes, Methods, and Pseudo Methods}
\label{sec:rprotobuf-classes}

@@

\section{Type Coercion}
\label{sec:types}

@@

\section{Evaluation: data.frame to Protocol Buffer Serialization}
\label{sec:evaluation}

@@

\section{Application: Data Interchange in Web Services}
\label{sec:opencpu}

@@

\section{Application: Distributed Data Collection with MapReduce}
\label{sec:mapreduce}

The MapReduce programming model \citep{dean2008mapreduce} has emerged
in the last decade as a popular framework for processing large data
sets in parallel on large compute clusters.  Protocol Buffers
provide a convenient mechanism to send structured data between tasks
in a MapReduce cluster.  In particular, the large data sets in fields
such as particle physics and information processing are frequently
stored in binned or histogram form in order to reduce the data storage
requirements for later data analysis \citep{scott2009multivariate}.

In such environments, analysts may be interested in very rare
phenomenon or be dealing with highly skewed data sets or inflexible
raw data storage systems from which unbiased sampling is not feasible.
There are two common patterns for generating histograms of large data
sets in a single pass with MapReduce.  In the first method, each
mapper task generates a histogram over a subset of the data that it
has been assigned, serializes this histogram and sends it to one or
more reducer tasks which merge the intermediate histograms from the
mappers.

In the second method, illustrated in
Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
point to a bucket width and outputs that bucket as a key and '1' as a
value.  Reducers then sum up all of the values with the same key and
output to a data store.

In both methods, the mapper tasks must choose identical bucket
boundaries in advance if we are to construct the histogram in a single
pass, even though they are analyzing disjoint parts of the input set
that may cover different ranges.  The \pkg{HistogramTools} package
\citep{histogramtools} enhances \pkg{RProtoBuf} by providing a concise
schema for R histogram objects.  The histogram message type is
designed to be helpful if some of the Map or Reduce tasks are written
in R, or if those components are written in other languages and only
the resulting output histograms need to be manipulated in R.
